Hierarchical Time Series Forecasting in Emergency Medical Services

### 2023-11-10 manuscript/02-literature.qmd
Ran styler to obtain better code formatting

```{r}
#| label: tbl-literature
#| echo: false
#| cache: false
#| tbl-cap: "Summary of some studies on forecasting in Emergency Medical Services"
library(kableExtra)
readxl::read_xlsx("../literature_table/table.xlsx") |>
  knitr::kable(booktabs = T, linesep = "") |>
  kable_styling(latex_options = "scale_down", font_size =11) |>
  column_spec(3,"6em") |>
  column_spec(4,"4em") |>
  column_spec(5,"10em")  |>
  column_spec(6,"8em") |>
  landscape()
```

>>>>

```{r}
#| label: tbl-literature
#| echo: false
#| cache: false
#| tbl-cap: "Summary of some studies on forecasting in Emergency Medical Services"
library(kableExtra)
readxl::read_xlsx("../literature_table/table.xlsx") |>
  knitr::kable(booktabs = T, linesep = "") |>
  kable_styling(latex_options = "scale_down", font_size = 11) |>
  column_spec(3, "6em") |>
  column_spec(4, "4em") |>
  column_spec(5, "10em") |>
  column_spec(6, "8em") |>
  landscape()
```

### 2023-11-10 manuscript/03-experiment.qmd
Ran styler to obtain better code formatting

```{r}
#| label: tbl-hierarchy
#| cache: true
#| tbl-cap: "Number of time series in each level for the hierarchical & grouped structure of attended incidents"
agg_level <- tibble::tribble(
  ~Level, ~`Number of series`,
  "All country", 1,
  "Control", 3,
  "Health board", 7,
  "Priority", 3,
  "Priority * Control", 3*3,
  "Priority * Health board", 3*7,
  "Nature of incident", 35,
  "Nature of incident * Control", 3*35,
  "Nature of incident * Health board", 7*35,
  "Priority * Nature of incident", 3*35-1,
  "Control * Priority * Nature of incident", 306,
  "Control * Health board * Priority * Nature of incident (Bottom level)", 691,
  "Total", 1530
)
knitr::kable(agg_level, booktabs = TRUE, position = "left") %>%
  kable_classic(full_width = FALSE) |> footnote(general = "Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset. As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.",threeparttable = TRUE)
```

@@

selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
p_nature <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & !is_aggregated(nature)) %>%
  as_tibble() %>%
  mutate(nature = as.character(nature)) %>%
  filter(nature %in% selected_nature) %>%
  group_by(date, nature, lhb) %>%
  summarise(incident = sum(incident), .groups = "drop") %>%
  ggplot(aes(x = date, y = incident, color = nature)) +
  geom_line() +
  labs(x = "Date", y = "Incidents", color = "Nature of incident")+
  ggthemes::scale_color_colorblind() +
  ggthemes::theme_few()

>>>>

```{r}
#| label: tbl-hierarchy
#| cache: true
#| tbl-cap: "Number of time series in each level for the hierarchical & grouped structure of attended incidents"
agg_level <- tibble::tribble(
  ~Level, ~`Number of series`,
  "All country", 1,
  "Control", 3,
  "Health board", 7,
  "Priority", 3,
  "Priority * Control", 3 * 3,
  "Priority * Health board", 3 * 7,
  "Nature of incident", 35,
  "Nature of incident * Control", 3 * 35,
  "Nature of incident * Health board", 7 * 35,
  "Priority * Nature of incident", 3 * 35 - 1,
  "Control * Priority * Nature of incident", 306,
  "Control * Health board * Priority * Nature of incident (Bottom level)", 691,
  "Total", 1530
)
knitr::kable(agg_level, booktabs = TRUE, position = "left") %>%
  kable_classic(full_width = FALSE) |>
  footnote(general = "Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset. As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.", threeparttable = TRUE)
```

@@

selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
p_nature <- incident_gthf %>%
  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & !is_aggregated(nature)) %>%
  as_tibble() %>%
  mutate(nature = as.character(nature)) %>%
  filter(nature %in% selected_nature) %>%
  group_by(date, nature, lhb) %>%
  summarise(incident = sum(incident), .groups = "drop") %>%
  ggplot(aes(x = date, y = incident, color = nature)) +
  geom_line() +
  labs(x = "Date", y = "Incidents", color = "Nature of incident") +
  ggthemes::scale_color_colorblind() +
  ggthemes::theme_few()

### 2023-11-10 manuscript/05-results.qmd
Ran styler to obtain better code formatting

```{r}
#| label: tbl-result
#| tbl-cap: "Average forecast performance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data. The test set consists of 462 days. The best approach is highlighted in bold. Point forecast accuracy is measured using MASE and MSSE, while probabilistic forecast accuracy is measured using CRPS."
#Set minimum in column to bold
set_to_bold <- function(table) {
  for(i in 3:6) {
    best <- which.min(table[[i]])
    table[[i]] <- sprintf(table[[i]], fmt="%1.3f")
    table[[i]][best] <- paste0("\\textbf{",table[[i]][best],"}")
  }
  return(table)
}

msse1 <- rmsse |>
  filter(
    method %in% c("mint", "base"),
    model != "qcomb",
    h > 42,
    model != "ensemble2"
  ) |>
  mutate(model = factor(model,
    levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
    labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT")),
  ) |>
  group_by(method, model, series) |>
  summarise(msse = mean(msse), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = msse) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
mase1 <- mase |>
  filter(
    method %in% c("mint", "base"),
    model != "qcomb",
    h > 42,
    model != "ensemble2"
  ) |>
  mutate(
    model = factor(model,
      levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
      labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
  ) |>
  group_by(method, model, series) |>
  summarise(mase = mean(mase), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = mase) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
crps1 <- readr::read_rds(here::here("results/crps.rds")) |>
  filter(method %in% c("mint", "base"),
         model != "qcomb",
         h > 42,
         model != "ensemble2"
  ) |>
  mutate(model = factor(model,
    levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
    labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
  ) |>
  group_by(method, model, series) |>
  summarise(crps = mean(crps), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = crps) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
bind_rows(mase1, msse1, crps1) |> 
  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "llrrrr") |> 
  pack_rows("MSSE", 1, 10) |> 
  pack_rows("MASE", 11, 20, hline_before = TRUE) |>
  pack_rows("CRPS", 21, 30, hline_before = TRUE) 
```

@@

# Plot of average accuracy vs week for each method for Total
acc_summary <- accuracy |>
  filter(
    series %in% c("Total", "Control areas", "Health boards"),
    method == "mint", model != "qcomb"
  ) |>
  mutate(model = case_when(
    model == "naiveecdf" ~ "Stationary",
    model == "ets" ~  "ETS",
    model == "tscount" ~ "TSGLM",
    model == "iglm" ~ "GLM",
    model == "ensemble" ~ "Ensemble",
    model == "ensemble2" ~ "Ensemble2"
  )) |>
  mutate(
    measure = factor(measure, levels = c("mase","msse","crps"), labels = c("MASE","MSSE","CRPS")),
    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
    model = factor(model, levels = c("Stationary", "ETS", "TSGLM", "GLM", "Ensemble", "Ensemble2")),
    week = factor(trunc((h - 1) / 7) + 1)
  ) |>
  group_by(week, model, measure, series) |>
  summarise(accuracy = mean(accuracy), .groups = "drop")

@@

incident_cv <- incident_gthf |> filter(lhb == "PO",is_aggregated(category),is_aggregated(nature))
fit <- incident_cv %>%
  model(
    ets = ETS(sqrt(incident))
  )
set.seed(1)
fcst <- fit %>% generate(h=1, times = 5000, bootstrap = TRUE)
p1 <- ggplot(data=fcst,aes(x=.sim, y=..density..)) +
  geom_histogram(binwidth = 1, fill="#d3d3d3", color="black")+
  geom_vline(data=fcst,aes(xintercept = (mean(.sim))), color="black", linetype="dashed", linewidth=1)+
  scale_x_continuous(breaks = seq(25,70,5))+
  scale_y_continuous(labels=percent)+
  ggthemes::theme_few()+
    labs(y="Density", x = "Demand", title = "Probabilistic forecast of ambulance demand for 1 day ahead")
p1
```

>>>>

```{r}
#| label: tbl-result
#| tbl-cap: "Average forecast performance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data. The test set consists of 462 days. The best approach is highlighted in bold. Point forecast accuracy is measured using MASE and MSSE, while probabilistic forecast accuracy is measured using CRPS."
# Set minimum in column to bold
set_to_bold <- function(table) {
  for (i in 3:6) {
    best <- which.min(table[[i]])
    table[[i]] <- sprintf(table[[i]], fmt = "%1.3f")
    table[[i]][best] <- paste0("\\textbf{", table[[i]][best], "}")
  }
  return(table)
}

msse1 <- rmsse |>
  filter(
    method %in% c("mint", "base"),
    model != "qcomb",
    h > 42,
    model != "ensemble2"
  ) |>
  mutate(
    model = factor(model,
      levels = c("naiveecdf", "ets", "iglm", "tscount", "ensemble", "ensemble2"),
      labels = c("Stationary", "ETS", "GLM", "TSGLM", "Ensemble", "Ensemble2")
    ),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT")),
  ) |>
  group_by(method, model, series) |>
  summarise(msse = mean(msse), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = msse) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
mase1 <- mase |>
  filter(
    method %in% c("mint", "base"),
    model != "qcomb",
    h > 42,
    model != "ensemble2"
  ) |>
  mutate(
    model = factor(model,
      levels = c("naiveecdf", "ets", "iglm", "tscount", "ensemble", "ensemble2"),
      labels = c("Stationary", "ETS", "GLM", "TSGLM", "Ensemble", "Ensemble2")
    ),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
  ) |>
  group_by(method, model, series) |>
  summarise(mase = mean(mase), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = mase) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
crps1 <- readr::read_rds(here::here("results/crps.rds")) |>
  filter(
    method %in% c("mint", "base"),
    model != "qcomb",
    h > 42,
    model != "ensemble2"
  ) |>
  mutate(
    model = factor(model,
      levels = c("naiveecdf", "ets", "iglm", "tscount", "ensemble", "ensemble2"),
      labels = c("Stationary", "ETS", "GLM", "TSGLM", "Ensemble", "Ensemble2")
    ),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
  ) |>
  group_by(method, model, series) |>
  summarise(crps = mean(crps), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = crps) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
bind_rows(mase1, msse1, crps1) |>
  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "llrrrr") |>
  pack_rows("MSSE", 1, 10) |>
  pack_rows("MASE", 11, 20, hline_before = TRUE) |>
  pack_rows("CRPS", 21, 30, hline_before = TRUE)
```

@@

# Plot of average accuracy vs week for each method for Total
acc_summary <- accuracy |>
  filter(
    series %in% c("Total", "Control areas", "Health boards"),
    method == "mint", model != "qcomb"
  ) |>
  mutate(model = case_when(
    model == "naiveecdf" ~ "Stationary",
    model == "ets" ~ "ETS",
    model == "tscount" ~ "TSGLM",
    model == "iglm" ~ "GLM",
    model == "ensemble" ~ "Ensemble",
    model == "ensemble2" ~ "Ensemble2"
  )) |>
  mutate(
    measure = factor(measure, levels = c("mase", "msse", "crps"), labels = c("MASE", "MSSE", "CRPS")),
    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
    model = factor(model, levels = c("Stationary", "ETS", "TSGLM", "GLM", "Ensemble", "Ensemble2")),
    week = factor(trunc((h - 1) / 7) + 1)
  ) |>
  group_by(week, model, measure, series) |>
  summarise(accuracy = mean(accuracy), .groups = "drop")

@@

incident_cv <- incident_gthf |> filter(lhb == "PO", is_aggregated(category), is_aggregated(nature))
fit <- incident_cv %>%
  model(
    ets = ETS(sqrt(incident))
  )
set.seed(1)
fcst <- fit %>% generate(h = 1, times = 5000, bootstrap = TRUE)
p1 <- ggplot(data = fcst, aes(x = .sim, y = ..density..)) +
  geom_histogram(binwidth = 1, fill = "#d3d3d3", color = "black") +
  geom_vline(data = fcst, aes(xintercept = (mean(.sim))), color = "black", linetype = "dashed", linewidth = 1) +
  scale_x_continuous(breaks = seq(25, 70, 5)) +
  scale_y_continuous(labels = percent) +
  ggthemes::theme_few() +
  labs(y = "Density", x = "Demand", title = "Probabilistic forecast of ambulance demand for 1 day ahead")
p1
```

### 2023-11-10 manuscript/main.qmd
Ran styler to obtain better code formatting

```{r}
#| label: setup
#| context: setup
#| include: false
library(fpp3)
library(knitr)
library(kableExtra)
library(ggthemes)
library(patchwork)
library(scales)
# Color blind color palettes
options(
  ggplot2.discrete.colour = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)
# Few theme
ggplot2::theme_set(ggthemes::theme_few())
```

```{r intro, child="01-introduction.qmd"}

```

```{r literature, child="02-literature.qmd"}

```

```{r experiment, child="03-experiment.qmd"}

```

```{r htc, child="04-htsgts.qmd"}

```

```{r results, child="05-results.qmd"}

```

```{r conclusion, child="06-conclusion.qmd"}

```

>>>>

```{r}
#| label: setup
#| context: setup
#| include: false
library(fpp3)
library(knitr)
library(kableExtra)
library(ggthemes)
library(patchwork)
library(scales)
# Color blind color palettes
options(
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)
# Few theme
ggplot2::theme_set(ggthemes::theme_few())
```

```{r intro, child="01-introduction.qmd"}
```

```{r literature, child="02-literature.qmd"}
```

```{r experiment, child="03-experiment.qmd"}
```

```{r htc, child="04-htsgts.qmd"}
```

```{r results, child="05-results.qmd"}
```

```{r conclusion, child="06-conclusion.qmd"}
```
